MCMC convergence diagnostics

### 2023-04-19 src/naomi-simple_mcmc/mcmc-convergence.Rmd
Add ESS plot to MCMC convergence notebook
MCMC convergence analysis for new run

---
title: "MCMC convergence diagnostics"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background** In the report `naomi-simple_fit` with parameter `tmbstan = TRUE`, we used the NUTS algorithm to perform MCMC inference for the simplified Naomi model.

@@

```{r}
out <- readRDS("depends/out.rds")
mcmc <- out$mcmc
```

@@

@@

We can also use energy plots (Betancourt 2017): ideally these two histograms would be the same 
When the histograms are quite different, it may suggest the chains are not fully exploring the tails of the target distribution.

>>>>

---
title: "MCMC convergence diagnostics"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
bibliography: citations.bib
abstract: |
  **Background** In the report `naomi-simple_fit` with parameter `tmbstan = TRUE`, we used the NUTS algorithm to perform MCMC inference for the simplified Naomi model.

@@

```{r}
out <- readRDS("depends/out.rds")
mcmc <- out$mcmc$stanfit
```

@@

# ESS

What are the total effective sample sizes?

```{r}
#' I think that this $summary should be all of the chains grouped together
mcmc_summary <- summary(mcmc)$summary

data.frame(mcmc_summary) %>%
  tibble::rownames_to_column("param") %>%
  ggplot(aes(x = n_eff)) +
    geom_histogram(alpha = 0.8) +
    labs(x = "ESS", y = "Count")
```

@@

We can also use energy plots [@betancourt2017conceptual]: ideally these two histograms would be the same 
When the histograms are quite different, it may suggest the chains are not fully exploring the tails of the target distribution.
